---
title: "Statistical Inference Course Project - Part 2"
output: 
  html_document:
    keep_md: yes
---

Prepared by Byron Stuart

## Overview

Basic inferential data analysis, analyze the ToothGrowth data in the R datasets package.
    
### 1. Load the ToothGrowth data and perform some basic exploratory data analyses

```{r, echo=TRUE}
data(ToothGrowth)
head(ToothGrowth)
str(ToothGrowth)

# Convert to factor for boxplot
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
```

### 2. Provide a basic summary of the data.

```{r, echo=TRUE}
summary(ToothGrowth)
library(ggplot2)
```

#### Further Analysis

```{r, echo=TRUE}
# t test impact on length by supplement
t.test(len ~ supp, ToothGrowth)
```

#### The p-value 0.06 is greater than an alpha of 0.05 (95% confidence interval) and the difference in the means spans 0 so we cannot say that one supplement is better than another

```{r, echo=TRUE}
graph <- ggplot(ToothGrowth, aes(x=supp, y=len)) + 
    geom_boxplot(aes(fill=supp)) +
    ylab("Length")  +
    xlab("Supp") +
    ggtitle("Length by Supplement")
graph
```

#### Based on averages alone OJ has longer lengths than VC but more investigation is required

```{r, echo=TRUE}
graph <- ggplot(ToothGrowth, aes(x=dose, y=len)) + 
    geom_boxplot(aes(fill=dose)) +
    ylab("Length")  +
    xlab("Dose") +
    ggtitle("Length by Dose")
graph
```

#### Considering both supplements the higher the dose the higher the average length is

### 3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.

#### 3a. Perform a two sample t test of OJ and VC

```{r, echo=TRUE}
t.test(len ~ supp, data=ToothGrowth, paired=FALSE, var.equal=FALSE)
```

#### The p-value 0.06 is greater than an alpha of 0.05 (95% confidence interval) and the difference in the means spans 0 hence the Null Hypothesis is "Failed to Reject", that is we cannot see a significant difference between supplements.

#### 3b. Perform a two sample t test at each dosage for OJ and VC (3 test results follow)

```{r, echo=TRUE}
t.test(len ~ supp, ToothGrowth[ToothGrowth$dose == .5, ])
t.test(len ~ supp, ToothGrowth[ToothGrowth$dose == 1, ])
```

#### At doses of 0.5 and 1 the p-value is less than an alpha of 0.05 (95% confidence interval) and the difference in the means does not span 0 hence the Null Hypothesis is "Rejected", that is we can see a significant difference between supplements.

```{r, echo=TRUE}
t.test(len ~ supp, ToothGrowth[ToothGrowth$dose == 2, ])
```

#### At a dose of 2 the p-value is greater than an alpha of 0.05 (95% confidence interval) and the difference in the means spans 0 hence the Null Hypothesis is "Failed to Reject", that is we cannot see a significant difference between supplements.

#### 3c. Perform separate two sample t test comparing doses of: 0.5 and 1; 1 and 2.

```{r, echo=TRUE}
t.test(ToothGrowth$len[ToothGrowth$dose==0.5], ToothGrowth$len[ToothGrowth$dose==1], paired=FALSE, var.equal=FALSE)
t.test(ToothGrowth$len[ToothGrowth$dose==1], ToothGrowth$len[ToothGrowth$dose==2], paired=FALSE, var.equal=FALSE)
```

#### For each of the 2 tests above the p-value is less than an alpha of 0.05 (95% confidence interval) and the difference in the means does not span 0. Hence the Null Hypothesis is "Rejected" in both cases showing there is a significant difference in length for higher doses of supplements. There is no need to test the difference between 0.5 and 2 given the above results.

### 4. State your conclusions and the assumptions needed for your conclusions.

#### Conclusions
1. Looking at all doses there is no significant difference in length for different supplements. This is based on the results of the t tests in section 3a.

2. Looking at both supplements there is significant difference in length for higher doses. This is based on the results of the t tests in section 3b.

3. Looking at the combined impact of supplement and dose there is significant difference in length for OJ compared to VC at a dose of 0.5mg and 1mg. At a dose of 2mg there is no significant difference in length between OJ and VC. In section 3c the t tests show that doses of 0.5mg and 1.0mg produce statistically significant p-values, indicating OJ outperforms VC. A dose of 2mg does not produce a statistically significant p-value indicating we cannot say whether OJ or VC are better.

#### Assumptions
1. The data follows a T distribution
2. The Data is derived from samples that are representative of the population
3. Variances between populations are different