---
title: "Statistical Inference Course Project - Part 1"
output: 
  html_document:
    keep_md: yes
---

Prepared by Byron Stuart

## Overview

Investigate the exponential distribution in R and compare it with the Central Limit Theorem (CLT). According to the CLT the distribution of sample means follow normal distribution.

Investigate the distribution of averages of 40 exponentials for a thousand simulations.

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials.

## Simulations

The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda.

```{r, echo=TRUE}
# Set the parameters
lambda <- 0.2
number <- 40
simulations <- 1000

# Generate 1000 simulations
exponentialDist <- replicate(simulations, mean(rexp(number, lambda)))
```

### Sample Mean versus Theoretical Mean

```{r, echo=TRUE}
# Calculate the simulated mean and the expected mean
mean <- mean(exponentialDist); mean
expectedMean <- 1/lambda; expectedMean

# Histogram of the simulated data
hist(exponentialDist, main="Mean Exponential Distribution", xlab="Means", breaks=40)

# Plot vertical line at the expected mean
abline(v=mean(exponentialDist), lwd="3", col="red")
text(4.5, 50, paste("Theoretical Mean =", expectedMean), pos = 2, col="red")
text(4.5, 45, paste("Mean =", round(mean,3)), pos = 2, col="blue")
```

#### Conclusion: The sample mean and theoretical mean are almost identical

### Sample Variance versus Theoretical Variance

```{r, echo=TRUE}
# Calculate the simulated standard error of the mean and compare it to the theoretical standard error
sd(exponentialDist); 1/(lambda*sqrt(number))

# Calculate the simulated variance of the mean and compare it to the theoretical variance
var(exponentialDist); 1/((lambda^2)*number)

# Compare the variance of 40 exponentials to a normal distribution
qqnorm(exponentialDist, col="blue")
qqline(exponentialDist, col=2)
legend("topleft", legend=c("Simulated", "Normal Distribution"), col=c("blue", "red"), pch=c(1, NA), lty=c(1,1), lwd=c(NA,3))
```

#### Conclusion: The sample variance and theoretical variance are almost identical

#### As per the CLT as the number of iterations increases the standard error and variance of the mean move closer to the theoretical values

### Distribution

```{r, echo=TRUE}
hist(exponentialDist, prob=TRUE, main="Mean Exponential Distribution", xlab="Means", breaks=40)
lines(density(exponentialDist), lwd=3, col="red")

# Add a Normal Curve
xfit<-seq(min(exponentialDist), max(exponentialDist), length=40)
yfit = dnorm(xfit, mean=1/lambda, sd=(1/lambda/sqrt(number)))
# yfit<-dnorm(xfit, mean=mean(exponentialDist), sd=sd(exponentialDist))
lines(xfit, yfit, col="blue", lwd=3)

legend("topright", legend=c("Means", "Normal Distribution"), col=c("red", "blue"), lty=c(1,1), lwd=c(3,3))
```

#### Conclusion: The distribution of the means closely matches a normal distribution, according to the central limit theorem (CLT) the distribution of averages of 40 exponentials is very close to a normal distribution.